### 2015-05-04 doc/201504_vnlbayes_params/slides_vnlb_prms.tex
Adds project documentation to repo

% \documentclass{beamer}
\documentclass[mathserif, 8pt]{beamer}

\mode<presentation> {
% \usetheme{Madrid}
%\usetheme{default}
\usetheme{Boadilla}
% \usecolortheme{beaver}
% \usecolortheme{rose}
}
% removes navigation symbols
\setbeamertemplate{navigation symbols}{}

\usepackage[english,activeacute]{babel}
\usepackage[english]{layout}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{pgf}
%\usepackage{movie15}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{color}
\usepackage[utf8]{inputenc}
\usepackage[multidot]{grffile}
\usepackage{ucs}
% \usepackage[T1]{fontenc}
%\usepackage{subfigure}

% from rida's preamble
\usepackage{epsfig,xspace}
\usepackage{setspace}
\usepackage{threeparttable}
\usepackage{subfloat}
\usepackage{pstricks,pst-node}
\usepackage[normal,tight,center]{subfigure}
\usepackage{pifont}
\usepackage{multimedia}
%\usepackage{media9}
%\usepackage{enumitem}
\usepackage{animate}
\usepackage{caption}
\usepackage{multirow}

% the following is used to bind frames with the same frame number
\usepackage{etoolbox}

\newcounter{multipleslide}

\makeatletter%
\newcommand{\multipleframe}{%
\setcounter{multipleslide}{\value{framenumber}}
\stepcounter{multipleslide}
\patchcmd{\beamer@@tmpl@footline}% <cmd>
	{\insertframenumber}% <search>
	{\themultipleslide}% <replace>
	{}% <success>
	{}% <failure>
}
\newcommand{\restoreframe}{%
\patchcmd{\beamer@@tmpl@footline}% <cmd>
	{\themultipleslide}% <search>
	{\insertframenumber}% <replace>
	{}% <success>
	{}% <failure>
\setcounter{framenumber}{\value{multipleslide}}%
}
\makeatother%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Running 'pdflatex --shell-escape...' will automatically recognize
% the eps files in the includegraphics and convert them to pdf.
% No file rename nor change to the document is needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{ifpdf}
\ifpdf
\DeclareGraphicsRule{.eps}{pdf}{.pdf}{`epstopdf #1}
\usepackage{epstopdf}
\epstopdfsetup{suffix=-\SourceExt-converted-to}
\pdfcompresslevel=0   %\pdfcompresslevel=9
\pdfcompresslevel0
\fi

% para poder incluir figuras .pdftex del xfig
\DeclareGraphicsRule{.pdftex}{pdf}{.pdftex}{}

\newcommand{\nada}[1]   {}

\newcommand{\best}[1]{\textbf{\textcolor{MyOrange}{#1}}}

\newcommand{\ma}[1]{\boldsymbol{#1}}
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\etal}{\textit{et al}. }
\newcommand{\tras}[1]{#1^{\mathrm{T}}}
\newcommand{\herm}[1]{#1^{\mathrm{H}}}
\newcommand{\con}[1]{#1^{\mathrm{*}}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\tech}[1]{\overline{#1}}
\newcommand{\nspace}{\!\!\!\!}
\newcommand{\nmbr}[1]{\oldstylenums{#1}}

\newcommand{\eps}{\varepsilon}
\newcommand{\R}{{\mathbb R}}
\newcommand{\Q}{{\mathbb Q}}
\newcommand{\N}{{\mathbb N}}
\newcommand{\Z}{{\mathbb Z}}
\newcommand{\C}{{\mathcal C}}
\renewcommand{\H}{{\mathcal H}}
\newcommand{\F}{{\mathcal F}}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\median}{median}

\definecolor{MyOrange}  {cmyk}{0,0.73,1.0,0}
\definecolor{MyGreen}   {cmyk}{1,0,1,0.0}

% Rida's colors and commands
\definecolor{Red}{rgb}{0.9,0.0,0.1}
\definecolor{Brown}{rgb}{0.55,0.27,0.1}
\definecolor{Brownie}{rgb}{0.75,0.27,0.1}
\definecolor{Yellow}{rgb}{1,1,0}
\definecolor{White}{rgb}{1,1,1}
\definecolor{formule}{rgb}{0.75,0.27,0.1}
\definecolor{formule}{rgb}{.7,1,0}
\definecolor{violet}{rgb}{0.7,0,.9}
\definecolor{darkgreen}{rgb}{0,.7,0}

%\newcommand{\reference}[1] {{\scriptsize \color{gray}  #1 }} 
\newcommand{\reference}[1] {{\color{gray}  #1 }} 
\newcommand{\rmk}[1]{{\color{Red} #1}}

%\theoremstyle{plain}\newtheorem{theorem}{Theorem}[chapter]
%\theoremstyle{plain}\newtheorem{proposition}{Proposition}[chapter]
%\theoremstyle{plain}\newtheorem{lemma}{Lemma}[chapter]
%\theoremstyle{definition}\newtheorem{definition}{Definition}[chapter]

% New definition of square root:
% it renames \sqrt as \oldsqrt
\let\oldsqrt\sqrt
% it defines the new \sqrt in terms of the old one
\def\sqrt{\mathpalette\DHLhksqrt}
\def\DHLhksqrt#1#2{%
\setbox0=\hbox{$#1\oldsqrt{#2\,}$}\dimen0=\ht0
\advance\dimen0-0.2\ht0
\setbox2=\hbox{\vrule height\ht0 depth -\dimen0}%
{\box0\lower0.4pt\box2}}

%gets rid of bottom navigation bars
% \setbeamertemplate{footline}[frame number]{}

%gets rid of navigation symbols
\setbeamertemplate{navigation symbols}{}

\title[VNLB - Parameter Analysis]{$\,$ \\ $\,$ \\ Parameter analysis for Video NL-Bayes}
\author[Pablo Arias]{March 31, 2015}
\institute[CMLA]{}
\date[]{}

%\author[Pablo Arias]{Pablo Arias$^*$, Vicent Caselles$^*$, Gabriele Facciolo$^\dagger$ \\ Rida Sadek$^*$, Guillermo Sapiro$^\ddagger$ \\ $\,$ \\  $\,$ \\ $\,$ \\ $\,$ \\
%\institute[UPF]{* Universitat Pompeu Fabra\\ $\dagger$ ENS Cachan \\ $\ddagger$ Duke University}

% \AtBeginSection[]  % "Beamer, do the following at the start of every section"
% {\begin{frame}<beamer>
% \frametitle{Outline} % make a frame titled "Outline"
% \tableofcontents[currentsection]  % show TOC and highlight current section
% \end{frame}}
%
% \AtBeginSubsection[]  % "Beamer, do the following at the start of every section"
% {\begin{frame}<beamer>
% \frametitle{Outline} % make a frame titled "Outline"
% \tableofcontents[currentsection,currentsubsection]  % show TOC and highlight current section
% \end{frame}}

\graphicspath{{./figures/}}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\section{Introduction}

\begin{frame}{Introduction}

	\begin{center}
	Experiments on 5 sequences of the Middlebury optical flow dataset:

	\bigskip

	\emph{Army}, \\
	\emph{DogDance}, \\
	\emph{Evergreen}, \\
	\emph{Mequon}, \\
	\emph{Walking},

	\bigskip

	with white additive Gaussian noise of

	\bigskip

	$\sigma = 10$, \\ 
	$\sigma = 25$, \\ 
	$\sigma = 40$.
	\end{center}

\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}

	Default parameters:

	\bigskip

	\begin{tabular}{l | c c | c c | c c }
		& \multicolumn{2}{c|}{$\sigma = 10$} 
		& \multicolumn{2}{c|}{$\sigma = 25$} 
		& \multicolumn{2}{c}{$\sigma = 40$} \\
		                            & 1st  & 2nd  & 1st  & 2nd  & 1st  & 2nd \\\hline\hline
		Patch size                  &  3   &   3  &  5   &   3  &  5   &   3 \\
		Number of patches           & 135  & 135  & 375  & 135  & 375  & 135 \\
		Distance threshold $\tau$   & n/a  & 432  & n/a  & 432  & n/a  & 432 \\
		Spatial search window       & 13   & 13   & 37   & 13   & 37   & 13  \\
		Temporal search range       & 5    & 5    & 5    & 5    & 5    & 5   \\\hline
		Beta                        & 1    & 1.2  & 1    & 1.2  & 1    & 1.2 \\\hline
	\end{tabular}

	\bigskip
	\bigskip
	\bigskip

	We consider varying the 2nd step spatial search window $W_x$ and patch size $p$.\\ All other parameters are kept constant.

	\end{center}
\end{frame}

\multipleframe
\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s10_army}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s25_army}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s40_army}

		\bigskip

		Results for Army.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s10_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s25_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s40_dogdance}

		\bigskip

		Results for DogDance.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s10_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s25_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s40_evergreen}

		\bigskip

		Results for Evergreen.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s10_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s25_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s40_mequon}

		\bigskip

		Results for Mequon.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s10_walking}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s25_walking}%
		\includegraphics[width=.33\textwidth]{prms_step2/psnr_bars-final_s40_walking}

		\bigskip

		Results for Walking.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\restoreframe

\multipleframe
\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 13.68 &  40.34 & 156.42 & 427.04 \\
			 21 & 14.94 &  49.54 & 195.54 & 666.86 \\
			 29 & 17.72 &  61.52 & 226.98 & 774.20 \\
			 37 & 19.42 &  70.20 & 270.96 & 895.74 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 10$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 27.48 &  55.62 & 176.20 & 489.34  \\
			 21 & 29.82 &  67.06 & 220.26 & 735.90  \\
			 29 & 33.12 &  75.06 & 262.62 & 878.22  \\
			 37 & 35.64 &  87.16 & 303.82 & 1014.76 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 25$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 2: final estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 28.14 &  59.42 & 200.92 & 616.32  \\
			 21 & 29.24 &  69.86 & 260.32 & 897.90  \\
			 29 & 32.46 &  79.86 & 308.46 & 1080.72 \\
			 37 & 36.52 &  95.78 & 362.98 & 1256.66 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 40$.
	\end{center}
\end{frame}
\restoreframe

\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}

	Default parameters:

	\bigskip

	\begin{tabular}{l | c c | c c | c c }
		& \multicolumn{2}{c|}{$\sigma = 10$} 
		& \multicolumn{2}{c|}{$\sigma = 25$} 
		& \multicolumn{2}{c}{$\sigma = 40$} \\
		                            & 1st  & 2nd  & 1st  & 2nd  & 1st  & 2nd \\\hline\hline
		Patch size                  &  3   &   3  &  5   &   3  &  5   &   3 \\
		Number of patches           & 135  & 135  & 375  & 135  & 375  & 135 \\
		Distance threshold $\tau$   & n/a  & 432  & n/a  & 432  & n/a  & 432 \\
		Spatial search window       & 13   & 13   & 37   & 13   & 37   & 13  \\
		Temporal search range       & 5    & 5    & 5    & 5    & 5    & 5   \\\hline
		Beta                        & 1    & 1.2  & 1    & 1.2  & 1    & 1.2 \\\hline
	\end{tabular}

	\bigskip
	\bigskip
	\bigskip

	We consider varying the 1st step spatial search window $W_x$ and patch size $p$.\\
	Two patch sizes will be consider in the second step: 3 and 5.\\
	All other parameters are kept constant.

	\end{center}
\end{frame}

\multipleframe
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s10_army}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s25_army}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s40_army}

		\bigskip

		Results for Army and $p^{(2)} = 3$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s10_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s25_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s40_dogdance}

		\bigskip

		Results for DogDance and $p^{(2)} = 3$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s10_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s25_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s40_evergreen}

		\bigskip

		Results for Evergreen and $p^{(2)} = 3$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s10_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s25_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s40_mequon}

		\bigskip

		Results for Mequon and $p^{(2)} = 3$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s10_walking}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s25_walking}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps3/psnr_basic-vs-final_s40_walking}

		\bigskip

		Results for Walking and $p^{(2)} = 3$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s10_army}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s25_army}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s40_army}

		\bigskip

		Results for Army and $p^{(2)} = 5$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s10_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s25_dogdance}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s40_dogdance}

		\bigskip

		Results for DogDance and $p^{(2)} = 5$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s10_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s25_evergreen}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s40_evergreen}

		\bigskip

		Results for Evergreen and $p^{(2)} = 5$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s10_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s25_mequon}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s40_mequon}

		\bigskip

		Results for Mequon and $p^{(2)} = 5$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s10_walking}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s25_walking}%
		\includegraphics[width=.33\textwidth]{prms_step1_step2_ps5/psnr_basic-vs-final_s40_walking}

		\bigskip

		Results for Walking and $p^{(2)} = 5$.\\From left to right: $\sigma = 10, 25$ and $40$.
	\end{center}
\end{frame}
\restoreframe

\multipleframe
\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 40.68 &  44.38 &  67.16 & 127.86 \\
			 21 & 42.48 &  46.10 &  74.72 & 152.92 \\
			 29 & 42.26 &  47.82 &  83.90 & 177.02 \\
			 37 & 43.28 &  52.10 &  91.70 & 200.28 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 10$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 48.52 &  52.96 &  86.74 & 180.52 \\
			 21 & 48.92 &  56.26 &  96.10 & 206.70 \\
			 29 & 49.40 &  57.12 & 100.72 & 216.30 \\
			 37 & 49.88 &  60.14 & 105.68 & 229.04 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 25$.
	\end{center}
\end{frame}

\begin{frame}{Parameters for step 1: basic estimate}
	\begin{center}
		\begin{tabular}{l | c c c c }
			 $W_x$\textbackslash $p$ & 3  & 5  & 7  & 9 \\\hline\hline
			 13 & 53.44 &  56.12 &  92.62 & 190.76 \\
			 21 & 54.00 &  59.84 & 101.28 & 222.16 \\
			 29 & 54.90 &  62.36 & 107.18 & 235.28 \\
			 37 & 55.00 &  63.00 & 111.58 & 250.36 \\
		\end{tabular}

		\bigskip

		Average running times (s) accross sequences for $\sigma = 40$.
	\end{center}
\end{frame}
\restoreframe

\multipleframe
\begin{frame}{PSNRs of basic estimates}

	\begin{center}
	\begin{tabular}{| c | c |c c c c c|}
		\hline \hline
		$\sigma$  & Method \textbackslash Video & Army & DogDance & Evergreen & Mequon & Walking \\\hline\hline
		\multirow{5}{*}{$10$} & BM4D           & \best{37.48} & \best{35.60} & \best{35.29} & \best{37.89} & \best{38.67} \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       14.30  &       18.98  &       15.87  &       16.83  &       16.28  \\%\cline{2-7}
		                      & VNLB $W_t = 1$ &       36.21  &       34.86  &       34.50  &       37.16  &       37.54  \\%\cline{2-7}
		                      & VNLB $W_t = 2$ &       36.77  &       35.22  &       34.82  &       37.69  &       38.20  \\\hline
%									 
		\multirow{5}{*}{$25$} & BM4D           & \best{32.24} & \best{31.07} & \best{30.48} &       32.89  & \best{33.54} \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       30.13  &       29.51  &       29.07  &       30.91  &       30.00  \\%\cline{2-7}
		                      & VNLB $W_t = 1$ &       31.92  &       30.85  &       30.33  &       32.86  &       33.00  \\%\cline{2-7}
									 & VNLB $W_t = 2$ & \best{32.21} & \best{31.02} & \best{30.50} & \best{33.14} &       33.31  \\\hline
%									 
		\multirow{5}{*}{$40$} & BM4D           & \best{29.53} & \best{28.59} &       27.93  &       29.94  & \best{30.55} \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       27.03  &       26.62  &       26.27  &       27.59  &       27.65  \\%\cline{2-7}
		                      & VNLB $W_t = 1$ &       29.13  &       28.30  &       27.88  &       29.79  &       29.91  \\%\cline{2-7}
									 & VNLB $W_t = 2$ & \best{29.50} & \best{28.55} & \best{28.13} & \best{30.15} &       30.29  \\\hline\hline
	\end{tabular}

	\bigskip
	Results of VNLB corresponding to patch sizes of $5$ in both steps.\\The rest of parameters are the default ones.
	\end{center}

\end{frame}

\begin{frame}{PSNRs of final estimates}

	\begin{center}
	\begin{tabular}{| c | c |c c c c c|}
		\hline \hline
		$\sigma$  & Method \textbackslash Video & Army & DogDance & Evergreen & Mequon & Walking \\\hline\hline
		\multirow{5}{*}{$10$} & BM4D           & \best{37.77} &       35.70  & \best{35.40} &       38.09  &       38.85  \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       34.46  &       33.61  &       33.05  &       35.65  &       35.90  \\%\cline{2-7}
		                      & VNLB $W_t = 1$ &       37.65  & \best{35.87} & \best{35.42} & \best{39.10} & \best{39.70} \\%\cline{2-7}
									 & VNLB $W_t = 2$ & \best{37.75} & \best{35.92} & \best{35.47} & \best{39.06} & \best{39.79} \\\hline
%									 
		\multirow{5}{*}{$25$} & BM4D           &       32.91  &       31.50  &       30.94  &       33.60  &       34.27  \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       32.33  &       31.17  &       30.39  &       33.69  &       33.74  \\%\cline{2-7}
									 & VNLB $W_t = 1$ &       33.48  & \best{31.98} & \best{31.22} & \best{35.10} &       35.25  \\%\cline{2-7}
		                      & VNLB $W_t = 2$ & \best{33.58} & \best{31.98} & \best{31.22} & \best{35.08} & \best{35.38} \\\hline
%									 
		\multirow{5}{*}{$40$} & BM4D           &       30.42  &       29.29  &       28.65  &       31.02  &       31.64  \\%\cline{2-7}
		                      & VNLB $W_t = 0$ &       29.94  &       28.95  &       28.24  &       30.98  &       30.95  \\%\cline{2-7}
									 & VNLB $W_t = 1$ &       31.36  &       29.96  & \best{29.19} & \best{32.70} &       32.71  \\%\cline{2-7}
									 & VNLB $W_t = 2$ & \best{31.55} & \best{30.01} & \best{29.24} & \best{32.75} & \best{32.94} \\\hline\hline
	\end{tabular}

	\bigskip
	Results of VNLB corresponding to patch sizes of $5$ in both steps.\\The rest of parameters are the default ones.
	\end{center}

\end{frame}
\restoreframe

\begin{frame}{}
	\centering
   \vspace{1.7cm}
   \textcolor{MyOrange}{\textbf{\Huge{Merci.}}}\\
   \vspace{5cm}
\end{frame}

% 
%

\end{document}